Hybrid platforms and bargaining power

### 2024-05-09 src/paper/application.tex
Add proofs for main text

Let us now examine entry fees and fringe entry decisions in the benchmark model.
Recall that there is an infinity of potential fringe entrants looking to enter the market.
Therefore, total profits in equilibrium must be zero.
Combined with the  profit function, this gives the following expressions for the equilibrium number of fringe firms and the equilibrium size of the aggregate.
\begin{proposition}
    \label{prop:equilibrium_aggregate}
    If entry costs $I_F$ and $F_F$ are low enough, the equilibrium size of the aggregate is
    \begin{align}
        A = \mu \frac{V_F}{F_F + I_F}.
        \label{eq:aggregate_eq}
    \end{align}
    and the equilibrium number of fringe firms is
    \begin{align*}
        N_F = \frac{\mu}{F_F + I_F} - N_P \frac{V_P}{V_F} - \frac{1}{V_F}.
    \end{align*}
    Otherwise,  $N_F = 0$ and the equilibrium size of the aggregate is
    \begin{align*}
        A = N_P V_P + 1.
    \end{align*}
\end{proposition}
Note that in the first case (\cref{eq:aggregate_eq}), the size of the aggregate does not depend on the platforms' product variety or the platform's product value.
The intuition behind this is that per-firm fringe profits only depend on these factors indirectly, through the size of the aggregate.
Therefore, the zero profit condition for the fringe firms pins down the aggregate in equilibrium (as long as the fringe is feasible).
That is, an increase in $N_P$ will simply replace some fringe entrants, but the free entry condition will pin down the same aggregate regardless of the number of platform products.

@@

First, let us look at the case when the platform finds it optimal to operate in the hybrid regime.
Notice that the optimal entry fee does not depend on either the platform's product value or product variety (\cref{fig:entry_fee}).
This is because due to the lump-sum nature of the entry fee, the platform can extract all the surplus from the fringe firms, and thus chooses $F_F$ to maximize total industry profits.
The latter is essentially a function of the aggregate, minus the investment costs of the fringe firms.
As I have shown in \cref{prop:equilibrium_aggregate}, the aggregate is independent of the platform's product variety in the hybrid regime, and therefore the optimal entry fee is also independent of it.

@@

Another important corollary of \cref{prop:optimal_entry_fee} is that consumer welfare is also unaffected by the platform's product variety in the hybrid regime.
\begin{corollary}
    \begin{align*}
        N_F > 0 \implies \frac{\partial CS}{\partial N_P} = 0.
    \end{align*}
\end{corollary}
This result is a direct consequence of the facts that the aggregate is independent of the platform's product variety in the hybrid regime, and that consumer surplus is a monotone increasing function of the aggregate.

@@

\begin{proposition}
    Assume that the platform is operating in the hybrid regime. Then, the equilibrium number of fringe firms is decreasing in the number of the platform's products. Furthermore, this decrease is more than proportional to the increase in the platform's product variety (also taking into account product value).
    \begin{align*}
        N_F > 0 \implies \frac{\partial N_F}{\partial N_P} < -\frac{V_P}{V_F} < 0.
    \end{align*}
\end{proposition}

\begin{corollary}
    Assume that the platform is operating in the hybrid regime. Then the equilibrium aggregate and consumer surplus is decreasing in the number of the platform's products.
    \begin{align*}
        N_F > 0 \implies \frac{\partial A}{\partial N_P}, \frac{\partial CS}{\partial N_P} < 0.
    \end{align*}
\end{corollary}

@@

When the fringe is not viable, the platform's product variety has the same, mechanical effect on total aggregate and consumer surplus as in the benchmark model.
\begin{proposition}
    Assume that the platform is operating in the pure retail regime. Then the equilibrium aggregate and consumer surplus is increasing in the number of the platform's products.
    \begin{align*}
        N_F = 0 \implies \frac{\partial A}{\partial N_P}, \frac{\partial CS}{\partial N_P} > 0.
    \end{align*}
\end{proposition}
This set of results is in stark contrast to the benchmark model.
Instead of increased product variety weakly increasing consumer welfare, when entry fees are negotiated, it can actually decrease it if the platform stays in the hybrid regime.

@@

In the benchmark model, the platform having its own product variety has a weakly positive effect.
As long as the market is in the hybrid regime, and there is a non-zero number of fringe firms, it has no effect on the aggregate or consumer surplus.
When the fringe is no longer feasible, and the platform is in the pure retail regime, the platform's product variety mechanically increases both.
\begin{theorem}
    In the benchmark model with unilateral price setting, in a hybrid regime equilibrium (with $N_F > 0$), the following results hold:
    \begin{itemize}
        \item The equilibrium number of fringe firms decreases in the number of the platform's products: $\frac{\partial N_F}{\partial N_P} = -\frac{V_P}{V_F} 0$.
        \item The aggregate is independent of the platform's product variety: $\frac{\partial A}{\partial N_P} = 0$.
        \item Consumer surplus is independent of the platform's product variety: $\frac{\partial CS}{\partial N_P} = 0$.
        % \item The platform's profit is increasing in the number of its products: $\frac{\partial \pi_P^t}{\partial N_P} = \frac{V_P}{V_F} I_F$.
    \end{itemize}
\end{theorem}

In contrast, in the bargaining model, increasing the platform's product variety can have a negative effect on consumer welfare.
The reason for this is that the platform's product variety increases its bargaining power, and thus decreases the share that the fringe firms obtain.
This in turn leads to a decrease in the number of fringe firms, and this decrease is large enough to offset the increase in the platform's product variety.
\begin{theorem}
    In the bargaining model, in a hybrid regime equilibrium (with $N_F > 0$), the following results hold:
    \begin{itemize}
        \item The equilibrium number of fringe firms decreases in the number of the platform's products more than in the benchmark case: $\frac{\partial N_F}{\partial N_P} < -\frac{V_P}{V_F} 0$.
        \item The aggregate decreases in the platform's product variety: $\frac{\partial A}{\partial N_P} < 0$.
        \item Consumer surplus decreases in the platform's product variety: $\frac{\partial CS}{\partial N_P} < 0$.
    \end{itemize}
\end{theorem}

@@

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has an additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.
\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let $f$ have the following, additive form:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$.
    Also assume that $g$ is twice differentiable and that $g'' < 0$.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
    Furthermore,
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_P} < 0.
    \end{align*}
\end{proposition}

@@

\begin{proof}[Proof of \cref{prop:equilibrium_aggregate}]
    From \cref{prop:optimal_profit}, the variable profit of each fringe firm is $\pi_{F_i}^{v*} = \mu V_F / A$.
    Total profit after entry fees and investment costs is therefore $\pi_{F_i}^{t*} = \mu V_F / A - I_F - F_F$.
    Under free entry, total profits are zero:
    \begin{align*}
        0 = \pi_{F_i}^{t*} = \mu V_F / A - F_F - I_F.
    \end{align*}
    Simple rearrangement gives the formula we are looking for,
    \begin{align*}
        A = \mu \frac{V_F}{F_F + I_F},
    \end{align*}
    and substituting in $A = N_P V_P + N_F V_F + 1$ yields the equilibrium number of fringe firms,
    \begin{align*}
        N_F = \frac{\mu}{F_F + I_F} - N_P \frac{V_P}{V_F} - \frac{1}{V_F}.
    \end{align*}
\end{proof}

@@

\begin{proof}[Proof of \cref{prop:unique_equilibrium}]
    First, consider the function
    \begin{align*}
        g(x) = \frac{sx}{1 + sx}.
    \end{align*}
    Also, let us define
    \begin{align*}
        G(x) &= \int_0^1 s g(sx) \ds \\
        &= \frac{\log(x+1)}{x} + \frac{1}{x(x+1)} - \frac{1}{x}.
    \end{align*}

@@

    Now let us consider the function $h(x) = N_C \mu g(N_P V_P + V_F x)$.
    By \cref{lem:single_crossing_affine_transformation}, $H(x) = \int_0^1 s h'(sx) \ds$ is also either concave or decreasing for any $x > 0$.
    Finally, notice that $H(x)$ is exactly the profit function of the fringe:
    \begin{align*}
        \pi_F(N_P, N_F) = \int_0^1 N_F h'(s N_F) \ds,
    \end{align*}
    thus \cref{ass:single_crossing} is satisfied, and in turn \cref{prop:unique_equilibrium_more_general} implies that the equilibrium number of fringe entrants is unique.
\end{proof}

>>>>

Let us now examine entry fees and fringe entry decisions in the benchmark model.
Recall that there is an infinity of potential fringe entrants looking to enter the market.
Therefore, total profits in equilibrium must be zero.
Combined with the  profit function, this gives the following expressions for the equilibrium number of fringe firms and the equilibrium size of the aggregate.
\begin{proposition}
    \label{prop:equilibrium_aggregate_benchmark}
    If entry costs $I_F$ and $F_F$ are low enough, the equilibrium size of the aggregate is
    \begin{align}
        A = \mu \frac{V_F}{F_F + I_F}.
        \label{eq:aggregate_eq}
    \end{align}
    and the equilibrium number of fringe firms is
    \begin{align*}
        N_F = \frac{\mu}{F_F + I_F} - N_P \frac{V_P}{V_F} - \frac{1}{V_F}.
    \end{align*}
    Otherwise,  $N_F = 0$ and the equilibrium size of the aggregate is
    \begin{align*}
        A = N_P V_P + 1.
    \end{align*}
\end{proposition}
Note that in the first case (\cref{eq:aggregate_eq}), the size of the aggregate does not depend on the platforms' product variety or the platform's product value.
The intuition behind this is that per-firm fringe profits only depend on these factors indirectly, through the size of the aggregate.
Therefore, the zero profit condition for the fringe firms pins down the aggregate in equilibrium (as long as the fringe is feasible).
That is, an increase in $N_P$ will simply replace some fringe entrants, but the free entry condition will pin down the same aggregate regardless of the number of platform products.

@@

First, let us look at the case when the platform finds it optimal to operate in the hybrid regime.
Notice that the optimal entry fee does not depend on either the platform's product value or product variety (\cref{fig:entry_fee}).
This is because due to the lump-sum nature of the entry fee, the platform can extract all the surplus from the fringe firms, and thus chooses $F_F$ to maximize total industry profits.
The latter is essentially a function of the aggregate, minus the investment costs of the fringe firms.
As I have shown in \cref{prop:equilibrium_aggregate_benchmark}, the aggregate is independent of the platform's product variety in the hybrid regime, and therefore the optimal entry fee is also independent of it.

@@

Another important corollary of \cref{prop:optimal_entry_fee} is that consumer welfare is also unaffected by the platform's product variety in the hybrid regime.
\begin{corollary}
    \label{prop:consumer_surplus_benchmark}
    \begin{align*}
        N_F > 0 \implies \frac{\partial CS}{\partial N_P} = 0.
    \end{align*}
\end{corollary}
This result is a direct consequence of the facts that the aggregate is independent of the platform's product variety in the hybrid regime, and that consumer surplus is a monotone increasing function of the aggregate.

@@

\begin{proposition}
    \label{prop:N_F_comparative_equilibrium}
    Assume that the platform is operating in the hybrid regime. Then, the equilibrium number of fringe firms is decreasing in the number of the platform's products. Furthermore, this decrease is more than proportional to the increase in the platform's product variety (also taking into account product value).
    \begin{align*}
        N_F > 0 \implies \frac{\partial N_F}{\partial N_P} < -\frac{V_P}{V_F} < 0.
    \end{align*}
\end{proposition}

\begin{corollary}
    \label{prop:aggregate_comparative_hybrid}
    Assume that the platform is operating in the hybrid regime. Then the equilibrium aggregate and consumer surplus is decreasing in the number of the platform's products.
    \begin{align*}
        N_F > 0 \implies \frac{\partial A}{\partial N_P}, \frac{\partial CS}{\partial N_P} < 0.
    \end{align*}
\end{corollary}

@@

When the fringe is not viable, the platform's product variety has the same, mechanical effect on total aggregate and consumer surplus as in the benchmark model.
\begin{proposition}
    \label{prop:aggregate_comparative_retail}
    Assume that the platform is operating in the pure retail regime. Then the equilibrium aggregate and consumer surplus is increasing in the number of the platform's products.
    \begin{align*}
        N_F = 0 \implies \frac{\partial A}{\partial N_P}, \frac{\partial CS}{\partial N_P} > 0.
    \end{align*}
\end{proposition}
This set of results is in stark contrast to the benchmark model.
Instead of increased product variety weakly increasing consumer welfare, when entry fees are negotiated, it can actually decrease it if the platform stays in the hybrid regime.

@@

In the benchmark model, the platform having its own product variety has a weakly positive effect.
As long as the market is in the hybrid regime, and there is a non-zero number of fringe firms, it has no effect on the aggregate or consumer surplus.
When the fringe is no longer feasible, and the platform is in the pure retail regime, the platform's product variety mechanically increases both.
\begin{theorem}
    \label{prop:benchmark_results}
    In the benchmark model with unilateral price setting, in a hybrid regime equilibrium (with $N_F > 0$), the following results hold:
    \begin{itemize}
        \item The equilibrium number of fringe firms decreases in the number of the platform's products: $\frac{\partial N_F}{\partial N_P} = -\frac{V_P}{V_F} < 0$.
        \item The aggregate is independent of the platform's product variety: $\frac{\partial A}{\partial N_P} = 0$.
        \item Consumer surplus is independent of the platform's product variety: $\frac{\partial CS}{\partial N_P} = 0$.
        % \item The platform's profit is increasing in the number of its products: $\frac{\partial \pi_P^t}{\partial N_P} = \frac{V_P}{V_F} I_F$.
    \end{itemize}
\end{theorem}

In contrast, in the bargaining model, increasing the platform's product variety can have a negative effect on consumer welfare.
The reason for this is that the platform's product variety increases its bargaining power, and thus decreases the share that the fringe firms obtain.
This in turn leads to a decrease in the number of fringe firms, and this decrease is large enough to offset the increase in the platform's product variety.
\begin{theorem}
    \label{prop:bargaining_results}
    In the bargaining model, in a hybrid regime equilibrium (with $N_F > 0$), the following results hold:
    \begin{itemize}
        \item The equilibrium number of fringe firms decreases in the number of the platform's products more than in the benchmark case: $\frac{\partial N_F}{\partial N_P} < -\frac{V_P}{V_F} 0$.
        \item The aggregate decreases in the platform's product variety: $\frac{\partial A}{\partial N_P} < 0$.
        \item Consumer surplus decreases in the platform's product variety: $\frac{\partial CS}{\partial N_P} < 0$.
    \end{itemize}
\end{theorem}

@@

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has an additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.
\begin{assumption}
    \label{ass:additive_profit}
    Assume that $f$ has the following, additive form in $N_P$ and $N_F$:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$, and $g$ is twice differentiable and that $g'' < 0$.
\end{assumption}

\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let $f$ have the additive form from \cref{ass:additive_profit}.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
    Furthermore,
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_P} < 0.
    \end{align*}
\end{proposition}

@@

\begin{proof}[Proof of \cref{prop:equilibrium_aggregate_benchmark}]
    From \cref{prop:optimal_profit}, the variable profit of each fringe firm is $\pi_{F_i}^{v*} = \mu V_F / A$.
    Total profit after entry fees and investment costs is therefore $\pi_{F_i}^{t*} = \mu V_F / A - I_F - F_F$.
    Under free entry, total profits are zero:
    \begin{align*}
        0 = \pi_{F_i}^{t*} = \mu V_F / A - F_F - I_F.
    \end{align*}
    Simple rearrangement gives the formula we are looking for,
    \begin{align*}
        A = \mu \frac{V_F}{F_F + I_F},
    \end{align*}
    and substituting in $A = N_P V_P + N_F V_F + 1$ yields the equilibrium number of fringe firms,
    \begin{align*}
        N_F = \frac{\mu}{F_F + I_F} - N_P \frac{V_P}{V_F} - \frac{1}{V_F}.
    \end{align*}
\end{proof}

@@

\begin{proof}[Proof of \cref{lem:shape_of_fringe_profit}]
    First, consider the function
    \begin{align*}
        g(x) = \frac{sx}{1 + sx}.
    \end{align*}
    Also, let us define
    \begin{align*}
        G(x) &= \int_0^1 s g(sx) \ds \\
        &= \frac{\log(x+1)}{x} + \frac{1}{x(x+1)} - \frac{1}{x}.
    \end{align*}

@@

    Now let us consider the function $h(x) = N_C \mu g(N_P V_P + V_F x)$.
    By \cref{lem:single_crossing_affine_transformation}, $H(x) = \int_0^1 s h'(sx) \ds$ is also either concave or decreasing for any $x > 0$.
    Finally, notice that $H(x)$ is exactly the profit function of the fringe:
    \begin{align*}
        \pi_F(N_P, N_F) = \int_0^1 N_F h'(s N_F) \ds,
    \end{align*}
    thus proving the lemma.
\end{proof}

\begin{proof}[Proof of \cref{prop:unique_equilibrium}]
    \Cref{lem:shape_of_fringe_profit} demonstrates that the fringe total profit function is either concave or decreasing in the number of fringe firms.
    Thus, \cref{ass:single_crossing} is satisfied, and \cref{prop:unique_equilibrium_more_general} implies that the equilibrium number of fringe entrants is unique.
\end{proof}

\begin{proof}[Proof of \cref{prop:fringe_profits_partial}]
    $\pi^t_F(N_P ,N_F)$ adheres to the \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing}.
    Furthermore, 
    \begin{align*}
        \frac{\partial^2 \pi^t_F(N_P ,N_F)}{\partial N_P \partial N_F} < 0 \quad \forall\, N_P, N_F \geq 0.
    \end{align*}
    Therefore, by \cref{prop:share_of_fringe},
    \begin{align*}
        \frac{\partial \pi^t_F(N_P, N_F)}{\partial N_P} < 0 \quad \forall\, N_P, N_F \geq 0.
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{prop:N_F_comparative_equilibrium}]
    $\pi^t_F(N_P ,N_F)$ adheres to the \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing,ass:additive_profit}.
    Therefore, by \cref{prop:aggregate_size_additive}
    \begin{align*}
        N_F > 0 \implies \frac{\partial N_F}{\partial N_P} < -\frac{V_P}{V_F}
    \end{align*}
    immediately follows.
\end{proof}

\begin{proof}[Proof of \cref{prop:aggregate_comparative_hybrid}]
    The first statement follows immediately from
    \begin{align*}
        A = N_P V_P + N_F V_F + 1
    \end{align*}
    and
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} < -\frac{V_P}{V_F}
    \end{align*}
    from \cref{prop:N_F_comparative_equilibrium}.
    The second part is a consequence of
    \begin{align*}
        CS \propto \log(A).
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{prop:aggregate_comparative_retail}]
    The first statement trivially follows from
    \begin{align*}
        A = N_P V_P + 1.
    \end{align*}
    The second part is a consequence of
    \begin{align*}
        CS \propto \log(A).
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{prop:benchmark_results}]
    The respective results are stated in \cref{prop:equilibrium_aggregate_benchmark} and \cref{prop:consumer_surplus_benchmark}.
\end{proof}

\begin{proof}[Proof of \cref{prop:bargaining}]
    The respective results are stated in \cref{prop:N_F_comparative_equilibrium} and \cref{prop:aggregate_comparative_hybrid}.
\end{proof}
